Classical Planning Algorithms ¡ on the Atari Video Games: Preliminary Results

### 2014-10-31 workshop-paper/paper.tex
Pass on Experimental Section
bib

\caption{\small 
\emph{Top:} Number of classical planning problems  and percentages solved by IW(1), IW(2), or neither. Instances obtained
by splitting benchmarks  with $N$ atomic goals into $N$ problems with atomic goals.
\emph{Bottom:} Number of instances solved by \IW\ in comparison with  Iterative Deepening (\ID),  
Breadth-First Search (\BRFS),   and Greedy Best First Search (\textit{GBFS}) guided by  additive heuristic.
Time and  memory outs  after 30 minutes and 2 GB resp. From \cite{nir:ecai2102}.
}
\label{table1}
\end{table}
%%%%
%%%

@@

IW is a purely exploration algorithm that does not take into account  the accumulated 
reward for selecting the next states to consider. As a simple, variant that combines
exploration and exploitation, we evaluated a \emph{best-first search} algorithm
with two queues: one queue  ordered first by  novelty measure 
(recall, novelty 1 means that the state is the first one to make some atom true),
and a second queue ordered by accumulated reward. In one iteration, the best first
search picks up the best node from one queue, and in the second iteration it picks
up the best node from the other queue. This way for combining multiple heuristics
is used in the LAMA planner \cite{richter:lama}, and was introduced in the
planner Fast Downward \cite{malte:fd,malte:alternating}. In addition, we break ties
in the first queue favoring states with largest accumulated reward, and 
in the second queue, favoring states with smallest novelty measure. Last, when
a node is expanded, it is removed from the queue, and its children are placed on 
both queues. The exception are the nodes with accumulated reward 0 that are only
placed in the first queue.
We refer to this best-first algorithm below, as 2BFS.

@@

The notion of width underlying the iterated width algorithm was developed in the context
of classical planning in order to understand why most of the hundreds of existing benchmarks
resulting from planning competitions appear to be relatively simple for current planners, even
though classical planning is PSPACE-complete \cite{bylander}. The partial answer is that 
most of these domains have a low width, and hence, can be solved in low polynomial time (by IW),
when goals contain a single atom. At the same time, most problems with multiple atomic goals are 
also easy too, as the goals can often be achieved one at a time \cite{nir:ecai2012}.

@@

We have shown experimentally that width-based algorithms like IW(1) and 2BFS
that originate in work in classical planning, can be used in the Atari games
with a performance that approaches the state-of-the-art as captured by UCT 
planning, sometimes using a fraction of the time. We are currently addressing three issues, 
which we haven't managed to get in time for this submission. First, we want to extend the evaluation to 
all the games featured in \cite{bellemare:jair}, and to add more runs per  game. 
Second, we want to evaluate the algorithm for different fixed time windows. In the table above, for example, UCT
is often taking an order-of-magnitude more time per decision as 2BFS, so the 
comparison in not fair. Third, we need to revise the way duplicate states are handled 
in IW and 2BFS to ensure that the best paths among the paths defined by the states that have been expanded
are returned. This change is important from a theoretical point of view; it's not clear
what it's practical impact will be.

>>>>

\caption{\small 
\emph{Top:} Number of classical planning problems  and percentages solved by IW(1), IW(2), or neither. Instances obtained
by splitting benchmarks  with $N$ atomic goals into $N$ problems with atomic goals.
\emph{Bottom:} Number of instances solved by \IW\ in comparison with  Iterative Deepening (\ID),  
Breadth-First Search (\BRFS),   and Greedy Best First Search (\textit{GBFS}) guided by  additive heuristic.
Time and  memory outs  after 30 minutes and 2 GB resp. From \cite{nir:ecai2012}.
}
\label{table1}
\end{table}
%%%%
%%%

@@

IW is a purely exploration algorithm that does not take into account  the accumulated 
reward for selecting the next states to consider. As a simple, variant that combines
exploration and exploitation, we evaluated a \emph{best-first search} algorithm
with two queues: one queue  ordered first by  novelty measure 
(recall, novelty 1 means that the state is the first one to make some atom true),
and a second queue ordered by accumulated reward. In one iteration, the best first
search picks up the best node from one queue, and in the second iteration it picks
up the best node from the other queue. This way for combining multiple heuristics
is used in the LAMA planner \cite{richter:lama}, and was introduced in the
planner Fast Downward \cite{helmert:fd,malte:alternating}. In addition, we break ties
in the first queue favoring states with largest accumulated reward, and 
in the second queue, favoring states with smallest novelty measure. Last, when
a node is expanded, it is removed from the queue, and its children are placed on 
both queues. The exception are the nodes with accumulated reward 0 that are only
placed in the first queue.
We refer to this best-first algorithm below, as 2BFS.

@@

The notion of width underlying the iterated width algorithm was developed in the context
of classical planning in order to understand why most of the hundreds of existing benchmarks
resulting from planning competitions appear to be relatively simple for current planners, even
though classical planning is PSPACE-complete \cite{bylander:complexity}. The partial answer is that 
most of these domains have a low width, and hence, can be solved in low polynomial time (by IW),
when goals contain a single atom. At the same time, most problems with multiple atomic goals are 
also easy too, as the goals can often be achieved one at a time \cite{nir:ecai2012}.

@@

We have shown experimentally that width-based algorithms like IW(1) and 2BFS
that originate in work in classical planning, can be used in the Atari games
with a performance that approaches the state-of-the-art as captured by UCT 
planning, sometimes using a fraction of the time. We are currently addressing three issues, 
which we haven't managed to get in time for this submission. First, we want to extend the evaluation to 
all the games featured in \cite{bellemare:jair2013}, and to add more runs per  game. 
Second, we want to evaluate the algorithm for different fixed time windows. In the table above, for example, UCT
is often taking an order-of-magnitude more time per decision as 2BFS, so the 
comparison in not fair. Third, we need to revise the way duplicate states are handled 
in IW and 2BFS to ensure that the best paths among the paths defined by the states that have been expanded
are returned. This change is important from a theoretical point of view; it's not clear
what it's practical impact will be.
